Effect of quality-adjusted production function inputs

### 2015-05-15 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Entire analysis is working again with efficient code.

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, data = historicalData[[country]], nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
}
  # Save models to disk.
  save(models, file=fileName)
}
@

<<Fitted_and_Residuals>>=
extract_fitted_and_resid <- function(mod, id){
  df <- mod$data[ , c("Year", "iGDP", "Country")]
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(mod)[c(1:length(df$Year))]
  df$Residuals <- resid(mod)[c(1:length(df$Year))]
  # Parse the id to find out whether this is quality-adjusted data
  id_list <- strsplit(id, split = "[.]")[[1]]
  if (id_list[[2]] == "unadj"){
    df$adjusted <- "Unadjusted"
  } else if (id_list[[2]] == "adj"){
    df$adjusted <- "Quality-adjusted"
  } else {
    stop("Unknown adjustment.")
  }
  if (id_list[[3]] == "noE"){
    df$energy <- "Without energy"
  } else if (id_list[[3]] == "withE"){
    df$energy <- "With energy"
  } else {
    stop("Unknown energy.")
  }
  df$model <- id_list[[4]]    
  return(df)
}
fitted_and_resid_data <- leaf_apply(models, extract_fitted_and_resid, class=c("CDEmodel", "cesEst", "LINEXmodel"))
# Create a data frame.
fitted_and_resid_data <- do.call(rbind, fitted_and_resid_data)

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$adjusted <- relevelFactor(as.factor(fitted_and_resid_data$adjusted), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With Energy"))
@

@@

<<Fitted and Residual Graphs>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ adjusted) +
    xy_theme()
}

residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ adjusted) +
    xy_theme()
}
@

@@

<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESkle" | model=="CESkl"))
@

<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESkle" | model=="CESkl"))
@

<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESlek" | model=="CESkl"))
@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESlek" | model=="CESkl"))
@

<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESekl" | model=="CESkl"))
@

<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESekl" | model=="CESkl"))
@

@@

<<PT CES-kl, warning=FALSE>>=
naturalCoef(models$PT$unadj$noE$CESkl)
naturalCoef(models$PT$adj$noE$CESkl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals,
\Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the Solow residual.

@@

<<PT CES-kle unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESkle)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESkle)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESkle)$lambda} 
                                                                    & \Sexpr{naturalCoef(models$PT$adj$withE$CESkle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESkle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESkle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESlek)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESlek)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESlek)$lambda} & \Sexpr{naturalCoef(models$PT$adj$withE$CESle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESlek)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESlek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
After energy has been added, quality-adjustment 
has very little effect on the Solow residual,
creating a slight increase from 
\Sexpr{naturalCoef(models$PT$unadj$withE$CESlek)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CESlek)$lambda}.

@@

<<PT CES-ekl unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CESekl)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CESekl)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CESekl)$lambda} 
                                                                          & \Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESekl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the ($ek$)$l$ nesting, 
the Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).
Adding energy decreases the Solow residual 
(from \Sexpr{naturalCoef(models$PT$unadj$withE$CESekl)$lambda} to 
\Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda})
when enegy is included in the model.
But, adding energy slightly increases the Solow residual
(from \Sexpr{naturalCoef(models$PT$adj$noE$CESkl)$lambda} to \Sexpr{naturalCoef(models$PT$adj$withE$CESekl)$lambda})
when quality-adjusted inputs are used.

@@

<<UK CES-kl, warning=FALSE>>=
naturalCoef(models$UK$unadj$noE$CESkl)
naturalCoef(models$UK$adj$noE$CESkl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
\Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.

@@

<<UK CES-kle unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESkle)
@

@@

<<UK CES-kle quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESkle)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESkle)$lambda} 
                                                                & \Sexpr{naturalCoef(models$UK$adj$withE$CESkle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CESkle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CESkle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-lek unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESlek)
@

@@

<<UK CES-lek quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESlek)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESlek)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CESlek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CESlek)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CESlek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-ekl unadjusted>>=
naturalCoef(models$UK$unadj$withE$CESekl)
@

@@

<<UK CES-ekl quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CESekl)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CESkl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CESkl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CESekl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CESekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

>>>>

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, data = historicalData[[country]], nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
@

<<Fitted_and_Residuals>>=
fitted_and_resid_data <- leaf_apply(models, class=c("CDEmodel", "cesEst", "LINEXmodel"), function(mod, id){
  df <- mod$data[ , c("Year", "iGDP", "Country")]
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(mod)[c(1:length(df$Year))]
  df$Residuals <- resid(mod)[c(1:length(df$Year))]
  # Parse the id to find out whether this is quality-adjusted data
  id_list <- strsplit(id, split = "[.]")[[1]]
  if (id_list[[2]] == "unadjusted"){
    df$flavor <- "Unadjusted"
  } else if (id_list[[2]] == "adjusted"){
    df$flavor <- "Quality-adjusted"
  } else {
    stop("Unknown adjustment.")
  }
  if (id_list[[3]] == "noE"){
    df$energy <- "Without energy"
  } else if (id_list[[3]] == "withE"){
    df$energy <- "With energy"
  } else {
    stop("Unknown energy.")
  }
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}) %>% do.call(rbind, .)

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$adjusted <- relevelFactor(as.factor(fitted_and_resid_data$flavor), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With Energy"))
@

@@

<<Fitted and Residual Graphs>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    xy_theme()
}

residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    xy_theme()
}
@

@@

<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@

<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@

<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="lek" | nest=="kl")))
@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@

<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

@@

<<PT CES-kl, warning=FALSE>>=
naturalCoef(models$PT$unadj$noE$CES$kl)
naturalCoef(models$PT$adj$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals,
\Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the Solow residual.

@@

<<PT CES-kle unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$kle)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$kle)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$kle)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adj$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$lek)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$lek)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$lek)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$withE$CESle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$lek)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$lek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
After energy has been added, quality-adjustment 
has very little effect on the Solow residual,
creating a slight increase from 
\Sexpr{naturalCoef(models$PT$unadj$withE$CES$lek)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CES$lek)$lambda}.

@@

<<PT CES-ekl unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$ekl)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$ekl)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$ekl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$ekl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$ekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the ($ek$)$l$ nesting, 
the Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).
Adding energy decreases the Solow residual 
(from \Sexpr{naturalCoef(models$PT$unadj$withE$CES$ekl)$lambda} to 
\Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda})
when enegy is included in the model.
But, adding energy slightly increases the Solow residual
(from \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} to \Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda})
when quality-adjusted inputs are used.

@@

<<UK CES-kl, warning=FALSE>>=
naturalCoef(models$UK$unadj$noE$CES$kl)
naturalCoef(models$UK$adj$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
\Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.

@@

<<UK CES-kle unadjusted>>=
naturalCoef(models$UK$unadj$withE$CES$kle)
@

@@

<<UK CES-kle quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CES$kle)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CES$kle)$lambda} 
                                                                & \Sexpr{naturalCoef(models$UK$adj$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-lek unadjusted>>=
naturalCoef(models$UK$unadj$withE$CES$lek)
@

@@

<<UK CES-lek quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CES$lek)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CES$lek)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CES$lek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CES$lek)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CES$lek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-ekl unadjusted>>=
naturalCoef(models$UK$unadj$withE$CES$ekl)
@

@@

<<UK CES-ekl quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CES$ekl)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CES$ekl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CES$ekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
